---
title: "Project - Practical Machine Learning"
author: Yessika Labrador
output: html_document
---

## Libraries
These libraries are needed to downloading and managing the data, also to create different models:
```{r cache=FALSE, message=FALSE}
library(dplyr)
library(RCurl)
library(caret)
library(randomForest)
library(rpart)
```

# Loading Data
The data used in this study is data gathered from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. This data is available here:

- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r cache=TRUE}
trainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

fileTrain <- "/tmp/train.csv"
fileTest <- "/tmp/test.csv"

download.file(url=trainURL, destfile=fileTrain, method='curl')
download.file(url=testURL, destfile=fileTest, method='curl')

```

With the data downloaded, is time to load it on R
```{r cache=TRUE}
train <- read.csv(fileTrain)
test <- read.csv(fileTest)
```

Some features about data are: 

- The number of atributtes are ```r dim(train)[1] ```
- The name of the users are ```r unique(train$user_name) ```
- The values to predict are ```r unique(train$classe) ```

# Preprocess
For this study I'm going to delete some variables using some criterias:

```{r cache=TRUE}
# Initial size
dim(train)

# Deleting information variables 
train <- train %>% select(-c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window)) 

# Deleting variables with NA or empty values
max_percentage <- dim(train)[1] * 0.90
train <- train %>% select_if(colSums(!is.na(.)) > max_percentage, colSums(. == "") > max_percentage) 

# Removing zero covariates 
nsv <- nearZeroVar(train, saveMetrics=TRUE)
train <- train[ , !(nsv$nzv)]

# Final size
dim(train)
```

# Creating models
For this project I'm going to split over the date to create 2 sets: training and validation. And later I'm going to fit 2 models and compare them both.

```{r}
set.seed(1234)
inTrain <- createDataPartition(train$classe, p = 0.7, list = FALSE)
training <- train[inTrain,]
validation <- train[-inTrain,]

```

RandomForest Model

```{r}
rfModel <- randomForest(classe ~ ., data = training)
rfPred <- predict(rfModel, validation)
rfMatrix <- confusionMatrix(rfPred, validation$classe)
rfMatrix$table
rfMatrix$overall['Accuracy']
```

RandomTree Model

```{r}
rpartModel <- rpart(classe ~ ., data = training, method = "class")
rpartPred <- predict(rpartModel, validation, type="class")
rpartMatrix <- confusionMatrix(rpartPred, validation$classe)
rpartMatrix$table
rpartMatrix$overall['Accuracy']
```

Because tha accurancy for RandomForest is better than RandomTree, RandomForest is the selected classifier.

# Test Models

Finally, I'm going to make predictions for the testing set using the model fitted with RandomForest
```{r}
predict(rfModel, test)
```

# Conclusions
The model created provided satisfatory predictions for testing data



